---
title: Time Varying Regression Lab 1
output:
  html_document:
    toc: true
    toc_float: true
---

# Load the data

```{r load_data, fig.align = "center", fig.height = 4, fig.width = 8}
load("landings.RData")
landings$log.metric.tons = log(landings$metric.tons)
landings = subset(landings, Year <= 1987)
```

# Look at the data

```{r look_data}
landings
```

# Subset the data

```{r}
val <- "Sardine"
subset(landings, Species==val)
```

# Class of the data

```{r class_data}
class(data)
```

Because it is a data.frame, you can call the columns like so

```{r call_col}
landings$log.metric.tons
```

# Plot data with ggplot

```{r plot_data, fig.align = "center", fig.height = 4, fig.width = 8}
library(ggplot2)
ggplot(subset(landings, Species==val), aes(x=Year, y=log.metric.tons)) +
  geom_line()
```

# Plot data with base plot

```{r plot_data2, fig.align = "center", fig.height = 4, fig.width = 8}
dat <- subset(landings, Species==val)
plot(dat$Year, dat$log.metric.tons, type="l")
```

---

# Fit linear regression

First add `t` to the data.  This is Year minus first Year.

```{r}
landings$t = landings$Year-landings$Year[1]
```

Use the poly function

$$log(Sardine catch) = \alpha + \beta t + \beta_2 t^2 + \beta_3 t^3 + \beta_4 t^4 + e_t$$

```{r tvreg.sardine3}
model <- lm(log.metric.tons ~ poly(t,4), data=landings, subset=Species==val)
```

# Show summary of the fit

```{r}
summary(model)
```

# Fit a raw 4th order polynomial

Normally you should use `poly()`.

```{r}
dat = subset(landings, Species==val)
model <- lm(log.metric.tons ~ t + I(t^2) + I(t^3) + I(t^4), data=dat)
```


```{r}
summary(model)
```

# Show the coefficients


```{r}
coef(model)
```

# Show the fitted values

This is the model predicted values.

```{r}
fitted(model)
```

# Show the residuals

This is the difference between the data and the fitted line.

```{r}
residuals(model)
```

# Plot model fit over the data

```{r}
plot(dat$t, dat$log.metric.tons)
lines(dat$t, fitted(model))
```

or with ggplot

```{r}
fitted.df <- data.frame(t = dat$t, fitted=fitted(model))
ggplot(dat, aes(x=t, y=log.metric.tons)) +
  geom_point() +
  geom_line(color='red',data = fitted.df, aes(x=t, y=fitted))
```

# Test the residuals for independence

Here we will use the [Ljung-Box](https://en.wikipedia.org/wiki/Ljung–Box_test) test.

`fitdf` is the number of estimated parameters in your model.  Look at the coefficiets to determine the number of estimated parameters.

```{r}
coef(model)
```

For the 4th-order polynomial, it is 5.  It is probably best to use the default `lag`.

```{r}
x <- resid(model)
Box.test(x, type = "Ljung-Box", fitdf=5)
```

If the p-value is less than 0.05, it indicates support for temporally autocorrelated residuals, which means that your assumption of uncorrelated residuals is not supported.

# Problems

1. Fit a 1st order polynomial to the sardine data.  This is just $C_t = \alpha + \beta t + e_t$.

1. Try fitting a 4-th order time-varying regression for one of the other species.

```{r}
unique(landings$Species)
```

a. Fit a 4th order polynomial.
b. Do a `summary(model)` and evaluate what level of polynomial is supported.
c. Plot the data and the fit on top.